The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2022-11-01 paper.tex
Changes from meeting

\title{The yaq Project: \\ A Standard for Scientific Instrumentation}
\author{Kyle F. Sunden}
\author{Daniel D. Kohler}
\author{Kent A. Meyer}
\author{Peter L. Cruz Parilla}
\author{John C. Wright}
\author{Blaise J. Thompson}
\affiliation{
  University of Wisconsin--Madison
}
\email{blaise.thompson@wisc.edu}

@@

Modern instrumentation development often involves incorporation of many dissimilar hardware peripherals into a single unified instrument.
Increasing availability of modular hardware has brought greater instrument complexity to small research groups.
The flexibility promised to instrument makers by large hardware availability stretches the capability of traditional, monolithic orchestration software.
In many cases, a lack of software flexibility leads creative researchers to feel frustrated, unable to perform experiments they envision.
Herein we describe \yaq{}, a software project defining a new standardized way of communicating with diverse hardware peripherals.
\yaq{} encourages a highly modular approach to experimental software development, which is well suited to address the experimental flexibility needs of complex instruments.
\yaq{} is architected to overcome hardware communication barriers that are insurmountable with typical experimental software.
A large number of hardware peripherals are already supported with tooling available to create more.
The \yaq{} standard enables collaboration among multiple research groups, increasing code quality while lowering development effort.

@@

\yaq{} introduces a concept called ``Traits'' which are collections of related messages that are shared among multiple protocols.
Motors implement the ``has-position'' trait, which defines ``\texttt{set\_position}'', ``\texttt{get\_position}'', and ``\texttt{get\_units}''.
Sensors would implement the ``is-sensor'' trait which defines ``\texttt{get\_measured}'', ``\texttt{get\_channel\_names}'', and ``\texttt{get\_channel\_units}''.
Protocols which implement a trait must support all of the messages from the trait.
Importantly, specific protocols can also implement arbitrary additional messages that are not defined by any trait.

>>>>

\title{The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation}
\author{Kyle F. Sunden}
\author{Daniel D. Kohler}
\author{Kent A. Meyer}
\author{Peter L. Cruz Parrilla}
\author{John C. Wright}
\author{Blaise J. Thompson}
\affiliation{
  University of Wisconsin--Madison
}
\email{blaise.thompson@wisc.edu}

@@

Modern instrumentation development often involves incorporation of many dissimilar hardware peripherals into a single unified instrument.
Increasing availability of modular hardware has brought greater instrument complexity to small research groups.
%The flexibility promised to instrument makers by large hardware availability stretches the capability of traditional, monolithic orchestration software.
This complexity stretches the capability of traditional, monolithic orchestration software.
In many cases, a lack of software flexibility leads creative researchers to feel frustrated, unable to perform experiments they envision.
Herein we describe \yaq{}, a software project defining a new standardized way of communicating with diverse hardware peripherals.
\yaq{} encourages a highly modular approach to experimental software development, which is well suited to address the experimental flexibility needs of complex instruments.
\yaq{} is designed to overcome hardware communication barriers that are insurmountable with typical experimental software.
A large number of hardware peripherals are already supported, with tooling available to expand support.
The \yaq{} standard enables collaboration among multiple research groups, increasing code quality while lowering development effort.

@@

\yaq{} introduces a concept called ``Traits'' which are collections of related messages that are shared among multiple protocols.
Motors implement the ``\texttt{has-position}'' trait, which defines ``\texttt{set\_position}'', ``\texttt{get\_position}'', and ``\texttt{get\_units}''.
Sensors would implement the ``\texttt{is-sensor}'' trait which defines ``\texttt{get\_measured}'', ``\texttt{get\_channel\_names}'', and ``\texttt{get\_channel\_units}''.
Protocols which implement a trait must support all of the messages from the trait.
Importantly, specific protocols can also implement arbitrary additional messages that are not defined by any trait.
